\section{Attributes and Attribute-sets}

An {\tt ADCL\_Attribute} is an abstraction for a particular characteristic of a function/implementation. Each attribute is represented by the set of possible values for this characteristic.  An {\tt ADCL\_Attrset} is a collection of ADCL attributes.
In the following, we present the constructors and destructors for both ADCL objects.

\begin{verbatim}
int ADCL_Attribute_create ( int maxnvalues, int *array_of_values, 
			    ADCL_Attribute *attr );

subroutine ADCL_Attribute_create ( maxnvalues, array_of_values, attr, ierror )
integer maxnvalues, attr, ierr
integer array_of_values (*)
\end{verbatim}
with
\begin{itemize}
\item {\tt maxnvalues}(IN): number of possible values for this attribute
\item {\tt array\_of\_values}(IN): array of possible values for this attribute. The 
	value of an attribute is represented by an integer value. The values have to be 
	monotony increasing. However, they do not have to be contiguous.
\item {\tt attr}(OUT): handle to the attribute object
\end{itemize} 	


\begin{verbatim}
int ADCL_Attribute_free   ( ADCL_Attribute *attr );

subroutine ADCL_Attribute_free ( attr, ierror )
integer attr, ierror
\end{verbatim}
with
\begin{itemize}
\item {\tt attr}(INOUT): handle to the ADCL attribute to be freed. The handle has to be a valid attribute.
    	Upon successful completion, the handle will be set to {\tt ADCL\_ATTRIBUTE\_NULL}.
\end{itemize}



\begin{verbatim}
int ADCL_Attrset_create ( int maxnum, ADCL_Attribute *array_of_attributes, 
			    ADCL_Attrset *attrset );

subroutine ADCL_Attrset_create ( maxnum, array_of_attributes, attrset, ierror )
integer maxnum, attrset, ierror
integer array_of_attributes (*)
\end{verbatim}
with
\begin{itemize}
\item {\tt maxnum}(IN): number of attributes to be grouped
\item {\tt array\_of\_attributes}(IN): array of attributes. Each entry of the array has to be a 
	valid ADCL attribute created with {\tt ADCL\_Attribute\_create}.
\item {\tt attrset}(OUT): handle to the attribute-set object
\end{itemize}

\begin{verbatim}
int ADCL_Attrset_free ( ADCL_Attrset *attrset );

subroutine ADCL_Attrset_free ( attrset, ierror )
integer attrset, ierror
\end{verbatim}
with
\begin{itemize}
\item {\tt attrset}(INOUT): handle to the ADCL attribute-set to be freed. The handle has to be a valid attribute-set.
    	Upon successful completion, the handle will be set to {\tt ADCL\_ATTRSET\_NULL}.
\end{itemize}