\section{Environmental control functions}

This section discusses the general functions required to establish the ADCL environment and to shut it down.
All ADCL functions return error codes, it is up to the application to take the appropriate actions, the library does not abort in case of an internal error such as system exhaustion. The only exception to that rule is if an error occurs within an MPI function called by ADCL, since MPI's default error behavior is to abort in case of an error. In case the user would like to have an MPI error code in this case, the default error handler to {\tt MPI\_COMM\_WORLD} has to be set to {\tt MPI\_ERRORS\_RETURN}. See also section x.z in the MPI-1 specification.

ADCL provides right now C and F90 interfaces of most of its functions. Similarly to the way MPI defines its interfaces, the fortran interface of a particular routines contains an additional argument, namely the error code. Furthermore, all ADCL objects are defined as integers in Fortran, leaning again on the approach chosen by MPI. A C application has to include the ADCL header file called {\tt ADCL.h}, a Fortran application has to include the file {\tt ADCL.inc} in any routine utilizing ADCL functions.

\subsection{Initializing ADCL}
\begin{verbatim}

int ADCL_Init ( void );

subroutine ADCL_Init ( ierror )
integer ierror

\end{verbatim}

{\tt ADCL\_Init} initializes the ADCL execution environment. The function allocates internal data structures required for ADCL, and has to be therefore called before any other ADCL function. Upon success, ADCL returns {\tt ADCL\_SUCCESS}. It is recommended to call {\tt ADCL\_Init) right after {\tt MPI\_Init}. It is erroneous to call {\tt ADCL\_Init} multiple times.

\subsection{Shutting down ADCL}

\begin{verbatim}

int ADCL_Finalize ( void );

subroutine ADCL_Finalize ( ierror )
integer ierror
\end{verbatim}

{\tt ADCL\_Finalize} deallocates internal data structures. Since the function deallocates internal structures and since the function does not check for ongoing communication, the function should be called at the very end of the application, but before {\tt MPI\_Finalize}. It is erroneous to call {\tt ADCL\_Finalize} multiple times.

\subsection{ADCL program skeletons}

Using the two functions described above, the following presents the skeleton for any ADCL application.
\begin{verbatim}
#include <stdio.h>
#include "mpi.h"
#include "ADCL.h"

int main ( int argc, char **argv ) 
{
   MPI_Init ( &argc, &argv );
   ADCL_Init ();
   
   ...
   ADCL_Finalize ();
   MPI_Finalize ();
   return 0;
}
\end{verbatim}

Accordingly, the fortran skeleton looks as follows:
\begin{verbatim}
program ADCLskeleton
   include 'mpif.h'
   include 'adcl.inc'
   
   integer ierror
   
   call MPI_Init ( ierror )
   call ADCL_Init (ierror )
   
   ...
   call ADCL_Finalize ( ierror )
   call MPI_Finalize ( ierror )
end program ADCLskeleton
\end{verbatim}