\section{Introduction and Terminology}

ADCL is an application level communication library aiming at providing the
highest possible performance for application level communication patterns, The
library provides for each communication pattern a large number of
implementations and incorporates a runtime selection logic in order to choose
the implementation leading to the highest performance of the application on
the current platform. The library provides multiple different runtime
selection algorithms, including a brute force search strategy which tests all available
implementations of a given communication pattern; a heuristic
relying on orthogonal tuning of attributes characterizing an implementation;
and a 2k factorial design search strategy designed for large seach spaces and
correlated attributes values.

ADCL is not supposed to be a replacement for MPI, but on add-on library. In
fact, ADCL relies and exploits many of the features of MPI, and does not offer
for example point-to-point transfer primitives. The ADCL API
offers high level interfaces of application level collective operations. The
high level interfaces are required in order to be able to switch within the
library the implementation of the according collective operation without
modifying the application itself. Thus, ADCL complements functionality
available within MPI. The main objects within the ADCL API are:


\begin{itemize}

\item {\tt ADCL\_Attribute}s, which are an abstraction for particular
  characteristic of a function/implementation. Each attribute is represented
  by the set of possible values for this characteristic.

\item {\tt ADCL\_Attrset}s, which represent a collection of ADCL attributes

\item {\tt ADCL\_Function}s, each of them being equivalent to an actual
  implementation of a particular communication pattern. An ADCL Function can
  have an attribute-set attached to it, in which case the values for each of
  the attributes in the attribute-set for this particular function have to be
  defined.

\item {\tt ADCL\_Fnctset}s, which represent a collection of ADCL functions
  providing the same functionality. All Functions in a function-set have to
  have the same attribute-set. ADCL provides pre-defined function sets, such
  as for neighborhood communication ({\tt ADCL\_FNCTSET\_NEIGHBORHOOD}) or for
  shift operations ({\tt ADCL\_FNCTSET\_SHIFT}). The user can however also
  register its own functions in order to utilize the ADCL runtime selection
  logic.

\item {\tt ADCL\_Topology} objects, which provides a description of the
  process topology and neighborhood relations within the application.


\item {\tt ADCL\_Vector} which specify the data structures to be used during
  the communication and the actual data. The user can for example {\it
    register} a data structure such as a vector or a matrix with the ADCL
  library, detailing how many dimensions the object has, the extent of each
  dimension, which parts of the matrix shall be used for communication, the
  basic datatype of the object, and the pointer to the data array of the
  object.

\item {\tt ADCL\_Vmap}: which defines portion of a vector to be used in a
  particular communication operation/


\item {\tt ADCL\_Request} objects, which combines a process topology, a
  function-set and a vector object. The application can initiate a
  communication by starting a particular ADCL request.

\item{\tt ADCL\_Timer} object which allows to register a request with a timer
  object in order to base the tuning of a functionset on a larger code section
  instead of the function execution time itself. In addition, the timer object
  allows to co-tune multiple functionsets within the same code sequence.

\end{itemize}

This document discusses each of the objects in details and explains the user
level API functions. Starting from the Fall 2013 release of ADCL, a
higher-level API has also been released with provides MPI like interfaces for
many of the predefined operations supported by ADCL. In addition, ADCL
supports the notion of historic learning for some predefined function sets,
i.e. learning across multiple executions of the same operation.
