\section{Vectors and Vector-maps}

An {\tt ADCL\_Vector} combined with the {\tt ADCL\_Vmap} specifies the data
structures  to be used during the communication and the actual data. 

For the collectives defined in the MPI standard it is noticed that the
arguments of the parameters can be separated into three groups: information
concerning the data (buffer, data type), concerning the process topology
(communicator, root) and related to the communication pattern (element counts,
reduction operation, array of element counts or displacements). As an example,
for the {\tt MPI\_Bcast} interface, the information about the data consists of
the buffer and the data type, root and the MPI communicator comm give
information about the process topology and count is related to the
communication pattern. As each of the parameters of the MPI collectives falls
into one of these three groups, they form the basis of the ADCL objects:
information concerning the data is stored in the ADCL vector object,
information concerning the process topology in the ADCL topology object and
information related to the communication pattern becomes part of the new vmap
object. As of today, there are five different Vmap types defined in ADCL:'
\begin{itemize}
\item {\tt HALO}: specifies the number of halo-cells in each direction (hwidth).
\item {\tt ALL}: a value specifiying the number of elements used in the
  communication (count).
\item{\tt REDUCE}: a value specifying the number of elements used in the
  communication (count) along  with an operation (op) to be performed on the
  data.
\item{\tt LIST}: an array of values specifying the number of elements
  communicated to each process, with the element at position $i$ in the array
  used to communicate to rank $i$. In addition, an array of displacements has
  to be provided following the same rules as for the array of counts
  arguments.
\item{\tt INPLACE}: an abstract vmap object that can be used similarly to
  MPI\_INPLACE as an argument to some function sets.

\end{itemize}

The following are the (brief) interfaces to allocate each Vmap object. For the
sake of brevity we do not detail all arguments at this point.

\begin{verbatim}
int ADCL_Vmap_halo_allocate ( int hwidth, ADCL_Vmap *vec );
int ADCL_Vmap_list_allocate ( int size, int* rcnts, int* displ, 
                              ADCL_Vmap *vec );
int ADCL_Vmap_allreduce_allocate ( MPI_Op op, ADCL_Vmap *vec );
int ADCL_Vmap_reduce_allocate ( MPI_Op op, ADCL_Vmap *vec );
int ADCL_Vmap_alltoall_allocate ( int scnt, int rcnt, ADCL_Vmap *vec );
int ADCL_Vmap_all_allocate ( ADCL_Vmap *vec );
int ADCL_Vmap_inplace_allocate ( ADCL_Vmap *vec );
int ADCL_Vmap_free  ( ADCL_Vmap *vec );
\end{verbatim}

Similarly to all other objects, the Vmap object can be freed after utilizion
using 

\begin{verbatim}
int ADCL_Vmap_free ( ADCL_Vmap *vmap )
\end{verbatim}

ADCL distinguishes between allocating a
vector and registering a vector. The difference is, that in the first case,
the library allocates the memory for the object as specified by the user,
while in the second case the memory has to be allocated by the application.


\begin{verbatim}
int ADCL_Vector_allocate_generic ( int ndims, int *dims, int nc, 
                           ADCL_Vmap vmap, MPI_Datatype dat, 
                           void *data, ADCL_Vector *vec )

\end{verbatim}
with
\begin{itemize}
\item {\tt ndims}(IN): number of dimension of the data structure excluding the
  {\tt nc} argument detailed below.
\item {\tt dims}(IN): array of dimension {\tt ndims} containing the extent of
  each dimension including the halo-cells.
\item {\tt nc}(IN): number of elements per entry. Since many scientific codes
  have matrices where each entry of the matrix is a submatrix itself, this
  argument gives the possibility to specify the dimension of the submatrix.
\item {\tt vmap}(IN): vector map to be used for this vector.
\item {\tt dat}(IN): basic MPI datatype describing the data type of the matrix
\item {\tt data}(OUT): pointer to the buffer allocated. This pointer will be
  required for calculations within the user code, since the buffer has been
  allocated by ADCL. ADCL guarantees, that a contiguous memory location has
  been allocated for multi-dimensional arrays. Please note, that the buffer
  pointer is {\bf not} the beginning of the data array, but the pointer to the
  multi-dimensional matrix itself. As an example if {\tt ndims}=2 and the
  temporary variable used within ADCL to allocate the 2-D array is called {\tt
    tmp\_matrix}, then the buffer pointer returned in this argument is {\tt
    data = tmp\_matrix}, which is not equal to {\tt data =
    \&(tmp\_matrix[0][0])} in C!
\item {\tt vec}(OUT): handle to ADCL vector object.
\end{itemize}
There is no Fortran interface defined for this routine. For a discussion, why
no Fortran interface is provided for this routine, please refer to the
discussion of {\tt MPI\_Alloc\_mem} in the MPI-2~\cite{mpi2} specification
section 4.1.1.

\hspace{1cm}
\begin{verbatim}
int ADCL_Vector_register_generic ( int ndims, int *dims, int nc, 
                                   ADCL_Vmap vmap, MPI_Datatype dat, 
                                   void *data, ADCL_Vector *vec )

subroutine ADCL_Vector_register_generic ( ndims,  dims, nc, 
                vmap, dat, data, vec, ierror )
integer ndims, nc, vmap, dat, vec, ierror
integer dims(*)
TYPE data(*)
\end{verbatim}
with
\begin{itemize}
\item {\tt ndims}(IN): number of dimension of the data structure excluding the
  {\tt nc} argument detailed below.
\item {\tt dims}(IN): array of dimension {\tt ndims} containing the extent of
  each dimension including the halo-cells.
\item {\tt nc}(IN): number of elements per entry. Since many scientific codes
  have matrices where each entry of the matrix is a submatrix itself, this
  argument gives the possibility to specify the dimension of the submatrix.
\item {\tt vmap}(IN): variable describing which parts of the matrix can be
  used for communication. The Vmap object is typically the result of an {\tt
    ADCL\_Vmap\_allocate*} function as shown above.
\item {\tt dat}(IN): basic MPI datatype describing the data type of the matrix
\item {\tt data}(IN): pointer to the data array. Please note, that the buffer
  pointer has to be the pointer to the multi-dimensional matrix itself, and
  {\bf not} the beginning of the data array. As an example if {\tt ndims}=2
  and the variable used within application is defined as {\tt double
    tmp\_matrix[10][10]}, the buffer pointer passed to ADCL has to be {\tt
    tmp\_matrix}, which is not equal to {\tt \&(tmp\_matrix[0][0])} in C!
\item {\tt vec}(OUT): handle to ADCL vector object.
\end{itemize}

\hspace{1cm}

\begin{verbatim}
int ADCL_Vector_free ( ADCL_Vector *vec );
\end{verbatim}
with
\begin{itemize}
\item {\tt vec}(INOUT): handle to the vector object allocated with {\tt
    ADCL\_Vector\-\_allocate}. It is illegal to call this function with a
  vector object registered with {\tt ADCL\_Vector\_register}. Upon successful
  completion, the handle will be set to {\tt ADCL\_VECTOR\_NULL}.
\end{itemize}
There is no Fortran interface defined for this routine. For a discussion, why
no Fortran interface is provided for this routine, please refer to the
discussion of {\tt MPI\_Alloc\_mem} in the MPI-2~\cite{mpi2} specification
section 4.1.1.  \hspace{1cm}

\begin{verbatim}
int ADCL_Vector_deregister ( ADCL_Vector *vec );

subroutine ADCL_Vector_deregister ( vec, ierror )
integer vec, ierror
\end{verbatim}
with
\begin{itemize}
\item {\tt vec}(INOUT): handle to the vector object registered with {\tt
    ADCL\_Vector\-\_register}. It is illegal to call this function with a
  vector object allocated with {\tt ADCL\_Vector\_allocate}. Upon successful
  completion, the handle will be set to {\tt ADCL\_VECTOR\_NULL}.
\end{itemize}


The following example assumes, that each entry of this matrix contains of a
vector with five elements. 

\begin{verbatim}
#include <stdio.h>
#include "ADCL.h"
#include "mpi.h"

/* Dimensions of the data matrix per process */
#define DIM0  8
#define DIM1  4

int main ( int argc, char ** argv ) 
{
    int dims[2];
    int nc=5, hwidth=1;
    double ***data;
    ADCL_Vector vec;
    ADCL_Vmap vmap;    

    MPI_Init ( &argc, &argv );
    ADCL_Init ();
    
    dims[0] = DIM0 + 2*hwidth;
    dims[1] = DIM1 + 2*hwidth;
    ADCL_Vmap_halo_allocate ( hwidth, &vmap);
    ADCL_Vector_allocate_generic ( 2,  dims, nc, vmap,
        MPI_DOUBLE, &data, &vec );
    
    /* now you can access the elements of the vector in order to perform 
       computations, e.g */
    for ( i=1; i<DIM0; i++ ) {
       for ( j=1; j<DIM1; j++) {
         for ( k=0; k < nc; k++ ) {
           data[i][j][k] = ...
       }
    }
    
    ADCL_Vector_free ( &vec );        
    ADCL_Vmap_free ( &vmap );
    ADCL_Finalize ();
    MPI_Finalize ();
    return 0;
}
\end{verbatim}


